Refinement for open automata

### 2021-06-17 2021-Refinement/report.tex
Many small changes

% transitions between sections
% TODO: CITER LES CONFERENCES

@@

Refinement relations verifies that and implementation satisfies some properties with respect to a specification.
The open pNets model is based around a hierarchical composition of generalised automata called pLTS.
A properties a refinement can check on them is the non introduction of deadlocks.
A deadlock being an unintended state where no further action is possible.

@@

The contributions of this report are:
\begin{enumerate}
\item The formalisation of composition for open automata without requiring open pNets.
\item The translation of some properties of refinement relations from LTS to open automata.
\item The characterisation of a composition that doesn't introduce deadlocks.
\item The introduciton of several refinement relations for open automata that behave nicely with composition.
\end{enumerate}

This report is organised as follows:
We begin with Section \ref{sec:notations} by giving some notations used throughout this paper.
Then there are several sections dedicated to define all the interesting objects and properties:
In Section \ref{sec:def} we give a clean and up to date definition of open automata.
In Section \ref{sec:comp} we define their composition without requiring the use of open pNets.
In Section \ref{sec:proofelts} we adapt to open automata several standard properties from refinement relations on LTS.
After that Section \ref{sec:prelref} is dedicated to progressively build a refinement relation and explain the elements of this construction.
Section \ref{sec:refinement} introduces and analyses the refinement relation on open automata.
Finally Section \ref{sec:ccl} recapitulates the contribution and presents the work left.

@@

An instanciated transition of the automaton is transition produced by any well-formed substitution of the unbound variables of the transition minus the automaton variables.
\end{defi}
The intuition of an open automaton is a partially defined LTS with variables, guards on transitions, and parametrised actions. The LTS is partial in the sense that one cn plug an LTS in the hole and the surrounding LTS will react to the actions emitted by the sub-LTS put in the hole.
Initially the automaton is in the initial state with an extension on all the variables of the initial valuation.
In any configuration, it can perform instanciated transitions which source state is the current state and which guard is satisfiable in the current valuation if the holes emit the indicated actions.
An instanciated transition is a transition where every variable that is not an automaton variable is instantiated with a value.
By performing the instanciated transition, the automaton emits the indicated action and updates its configuration according to the target state and variable assignments.
\begin{noti}[FH-bisimulation]
The FH-bisimulation \cite{henrio:01055091} is noted \(\cong\).

@@

An run of this open automaton can be: The automaton in the hole \(l\) emits many actions, this synchronises with the only transition possible at that moment because \(v = L\).
Then the automaton in hole \(l\) emits a \(\delta\mpar{t}\), this is synchonised with the action that sets \(v \gets R\) and the automata in hole \(r\) must emit a \(\act{acc}\mpar{t}\).
Finally the automaton in the hole \(r\) emits any sequence of actions.

@@

\section{Composition of Open Automata}\label{sec:comp}
Open automata are partially specified automata, that partiality comes mostly from the holes.
The interpretation of a hole is an interface with another open automaton, in which we can plug an open automaton. The plugging operation is called composition.
In this report, composition will only refer to filling holes, other compositions like ``parallel composition" (pure interleaving) are not tackled (but one can define an open automaton that performs the parallel composition of two others).
The composition of open automata was already implicitly defined by the means of composition on pNets in previous work \cite{henrio:01299562} but never completely formalised on open automata.
The definition of composition below is a direct translation of what happens with pNets composition without the need of introducing pNets.
\begin{defi}[Composition of open automata]
The composition of \(A_c \defobject \OAg[c]\) in the hole \(k \in J_p\) of \(A_p \defobject \OAg[p]\) is an open automaton defined as follows.
\begin{align*}
	A_p\subst{A_c}{k} \defnotation & \OA{S_p \times S_c}{\mpar{s_{0p}, s_{0c}}}{V_p \uplus V_c}{\sigma_{0p} \uplus \sigma_{0c}}{J_c \uplus J_p \setminus \mbrc{k}}{T} \\
	\text{with } T \defobject & \mset{\OT{\mpar{s_p, s_c}}{\mpar{s'_p, s'_c}}{\alpha_p}{\beta_j^{j \in J'_c \uplus J'_p \setminus \mbrc{k}}}{g_p \wedge g_c \wedge \alpha_c = \beta_k}{\psi_p \uplus \psi_c}}{\OTx{p}{}{}{p} \in T_p, \OTx{c}{}{}{c} \in T_c} \\
	& \cup \mset{\OT{\mpar{s_p, s_c}}{\mpar{s'_p, s_c}}{\alpha_p}{\beta_j^{j \in J'_p}}{g_p}{\psi_p}}{\OTx{p}{}{}{p} \in T_p, k \notin J'_p, s_c \in S_c}
\end{align*}
\end{defi}
The action emitted when \(A_c\) makes a transition is sychronised with the action of the hole \(k\) in transitions of \(A_p\) which have it as a hole action (first transition set, \(\alpha_c = \beta_k\)).
The composition may look like a handshake, with both automata running in parallel (product of states and joint variables) however it is asymmetric because it is the first automaton that decides when the second can evolve, while the first automaton can evolve independently as soon as the hole is not involved in the open transition considered.

@@

\subsection{Properties of the relation on states}
The important properties we consider for the relation on configurations are being a pre-simulation and preventing the introduction of deadlocks.
They are defined in the following.
\begin{defi}[Pre-simulation]
Let \(A_1 \defobject \OAg[1]\) and \(A_2 \defobject \OAg[2]\).
A relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a pre-simulation if it satisfies both
\defitem Initial configurations are related: \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\);
\defitem From related states, all out-transitions from \(A_1\) can be simulated in \(A_2\) and their target states are related:
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}, \forall \sigma: V_1 \uplus V_2 \uplus \fvars{t_1} \to \values, \\
	\mpar{\sigma \vdash R\mpar{s_1, s_2} \wedge g_1} \implies \bigsymb{\exists} t_2 \defobject \OTx{2}{}{2}{2} \in \fOT{s_2}, \exists \nu: \fvars{t_2} \to \values, \\
	\sigma \uplus \nu \vdash \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \wedge g_2 \wedge R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}
\end{multline*}
\end{defi}
This definition translated to open automata the classical definition of simulation \cite{10.1007/3-540-54430-5_78}.
The last formula means that for every pair of related configurations of the automaton and every possible transitions (\(\vdash g_1\)) from the first automaton, there is a possible transition (\(\vdash g_2\)) such that the produced action matches (\(\alpha_1 = \alpha_2\)), the same holes' action (\(J'_1 \cap J'_2\)) matches (\(\beta_{1j} = \beta_{2j}\)) and the target states are related after variable update. % OR variableS
Apart from holes, this is a natural extension of the notion of simulation on LTS, which is the same definition without variables nor holes nor guards.
In FH-bisimulation, holes with the same name will receive the same open automaton when composed, and the resulting open automaton will still be equivalent.
Here we will assume that one of the related open automata can be obtained by composing an open automaton in a hole, open automaton which can have holes itself.
This implies that there is no relation between the related automata's holes, for instance composing \(A_i\) in \(j_i\) (\(1 \leq i \leq n\)) of \(A_0\) gives the set of holes \(J \defobject \mpar{\biguplus_{0 \leq i \leq n} J_i} \setminus \mset{j_i}{1 \leq i \leq n}\) that is neither subset nor superset of \(J_0\).
Yet we still assume that in holes with the same name, the same open automaton will be composed.
Hence only the holes in common have to match their actions.
However it is not sufficient to imply properties on the relation on automatons like being a preorder.
It is only a necessary condition that we put on any relation that we call a ``simulation".

In addition to the above property, we expect from a refinement relation to prevent the introduction of deadlocks.
Deadlocks are reachable configurations from which no further transition is possible.
They can arise unpredictably when parallelism is involved (which is the case in our composition) and they generally break the intended behaviour of a system.
Preventing them is also a required prerequisite for the refinement to preserve any liveness property.
\begin{defi}[Deadlock prevention, intuitive definition]\label{def:dpid}
A relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is deadlock reducing if:
\defitem It is a pre-simulation;
\defitem If from related states, there is a possible transition in the specification, then there is a pair of possible matching transitions from both automata:
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \forall \sigma: V_1 \uplus V_2 \to \values, \mpar{\sigma \vdash R\mpar{s_1, s_2}} \implies \\
	\mpar{\exists t_2 \in \fOT{s_2}, \exists \nu: \fvars{t_2} \to \values, \sigma \uplus \nu \vdash \fguard{t_2}} \implies \\
	\bigsymb{\exists} \mpar{t_1, t_2} \defobject \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2}, \exists \nu: \fvars{t_1} \uplus \fvars{t_2} \to \values, \\
	\sigma \uplus \nu \vdash g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \wedge R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}
\end{multline*}
\end{defi}
Actually, this requirement prevents the presence of deadlocks in the first automaton if there is a related non-deadlock configuration in the second.
The exact formulation is the contraposition of the one in the previous sentence.
Two reasons motivated the formulation given in Definition \ref{def:dpid}:
\begin{itemize}
\item We require a pair of matching transitions from both automata because we do not want a transition in the implementation making a configuration not a deadlock but this transition cannot be taken from the related configuration of the specification.
\item The alternative formulation of the property would have required the unintuitive quantification on a possibly empty set of transition to state that they are not satisfied.
\end{itemize}
Actually the first point is not necessary, the definition can be simplified into a less intuitive version:
\begin{defi}[Deadlock prevention, working definition]\label{def:dpwd}
A relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is deadlock reducing if it is a pre-simulation and it satisfies the following:
\[ \forall \mpar{s_1, s_2} \in S_1 \times S_2, V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \]
\end{defi}
The requirement here is that from any related configurations, if there is a possible transition in the second automaton (whatever the value of the free variables) then there is a possible transition in the first automaton (the existential quantifier is implicit in the notation \(\vdash\)).
The transition are not matched because the relation on configurations is a pre-simulation and it is sufficient to ensure it as will be shown in the proof.
Apart from the matching transitions this is exactly the same explaination as the intuitive definition.
The expected advantage of this formulation is that a SMT should behave better without the quantifiers.
Indeed the big disjunctions are not quantifiers because in practice the out-transitions are finitary, and so is the expansion of the disjunction.
\begin{lem}\label{lem:dred}
The intuitive (Definition \ref{def:dpid}) and working (Definition \ref{def:dpwd}) definitions of deadlock prevention are equivalent.
\end{lem}
Proof of the lemma is given in Appendix \ref{apx:lemdred}.

@@

For transitivity we need to introduce thee open automata.
Let \(A_1 \defobject \OAg[1]\), \(A_2 \defobject \OAg[2]\) and \(A_3 \defobject \OAg[3]\) be open automata.
Let \(R_{12}\) be a hole-\(\triangle\) simulation of \(A_1\) by \(A_2\) and \(R_{23}\) of \(A_2\) by \(A_3\).
The idea for \(R\) is that there exists a state \(s_2 \in S_2\) and values for \(V_2\) such that \(R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3}\).
However we cannot talk about states in formulas.
The solution is to expand the existential quantifier on states as we did in deadlock prevention, which is possible because states are finite.
\[ R \defobject \mpar{s_1, s_3} \mapsto \exists V_2, \bigvee_\subbox{s_2 \in S_2} R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3} \]
\(R\) is a witness of the transitivity, proof is given in \Ludo{Appendix \ref{apx:trans}}{are we supposed to read appendix D before C.2?}.
\end{proofsketch}
\begin{prop}[Limited composition correctness]\label{prop:cc'} Hole-subset simulation is correct w.r.t.\@ composition with automata without holes. \end{prop}
\begin{proofsketch}
For two open automata \(A_1 \defobject \OAg[1]\) and \(A_2 \defobject \OAg[2]\) with \(J_2 = \emptyset\), let \(k \in J_1\) and \(A \defobject A_1\subst{A_2}{k}\).
Let \(\reach{A}\) be a witness that \(A\) is non-locking.
The idea for the relation on configurations is that the copy of \(A_1\) in \(A\) do the same things as \(A_1\), as if we were trying to prove reflexivity.
\[ R \defobject \mpar{\mpar{s_{1'}, s_2}, s_1} \mapsto \choice{\reach{A}\mpar{s_{1'}, s_2} \wedge \bigwedge_\subbox{v_1 \in V_1} v_{1'} = v_1 & \text{if } s_{1'} = s_1 \\ \bot & \text{otherwise}} \]
Where \(y_{1'}\) is the renaming of \(y_1\) from \(A_1\) in \(A\).
\(R\) is a witness of the limited composition correctness, proof is given in Appendix \ref{apx:lchs}.
\end{proofsketch}
\begin{prop}[Special composition correctness]
If there is a hole-superset simulation of \(A_1\) by \(A_2\) and \(A_1\subst{A_3}{k}\) is non-locking, then there is a hole-superset simulation of \(A_1\subst{A_3}{k}\) by \(A_2\).
Where \(A_1 \defobject \OAg[1]\), \(A_2 \defobject \OAg[2]\), \(A_3\) are open automata and \(k \in J_1 \setminus J_2\).
\end{prop}
\begin{proofsketch}
Let \(R_{12}\) be a hole-superset simulation of \(A_1\) by \(A_2\) and \(\reach{A_1\subst{A_3}{k}}\) be a witness that \(A_1\subst{A_3}{k}\) is non-locking.
The idea for the relation on configurations is to combine the non-locking composition and the already existing realtion on configurations.
\[ R \defobject \mpar{\mpar{s_1, s_3}, s_2} \mapsto R_{12}\mpar{s_1, s_2} \wedge \reach{A_1\subst{A_3}{k}}\mpar{s_1, s_3} \]
\(R\) is a witness of the secial composition correctness, proof is given in Appendix \ref{apx:schs}.\Ludo{}{$y1$ et $y1'$ sortent de nulle part}
\end{proofsketch}
For composition correctness, the hole-subset simulation can only tell if there is a refinement when the composed automata have no holes.
To lift this limitation we need to allow adding holes which doesn't have a direct interpretation.
At first a solution using a function mapping the action performed by the new holes in the composition to the ones performed by removed hole in the base automaton was explored.
It was abandonned because the base automaton has nothing intelligent to tell about the action required by the composed automata.
That is the reason the hole-superset simulation doesn't constrain the action performed by the added holes.

\Ludo{Only one of the}{this paragraph is not clear} hole-\(\triangle\) simulations cannot tell that a non-locking composition with any automaton is a refinement because we need at the same time to add and remove holes, hence the need of the special composition correctness for hole-superset.
Also an issue arises when trying to tell that a composition is a refinement of the base automaton: the refinement relation cannot compare the holes sets.
This conflicts with the transitivity propriety when a hole present in both extrema of the chain is absent in an intermediate automaton.
In that case all the constraint on holes actions are lifted even if we would like to require an equality of these actions on the transitive cloture.

>>>>

@@

Refinement relations verifies that and implementation satisfies some properties with respect to a specification.
The open pNets model is based around a hierarchical composition of generalised automata.
A properties a refinement can check on them is the non introduction of deadlocks.
A deadlock being an unintended state where no further action is possible.

@@

The contributions of this report are:
\begin{enumerate}
\item The formalisation of composition for open automata without requiring open pNets.
\item The translation of some properties of refinement relations from LTS to open automata.
\item The characterisation of a composition that doesn't introduce deadlocks.
\item The introduciton of several refinement relations for open automata that behave nicely with composition.
\end{enumerate}
This report is organised as follows:
We begin with Section \ref{sec:sota} by describing the state-of-the-art.
Then there are several sections dedicated to define all the interesting objects and properties:
In section \ref{sec:notations} we give some notations used throughout this paper.
In Section \ref{sec:def} we give a clean and up to date definition of open automata.
In Section \ref{sec:comp} we define their composition without requiring the use of open pNets.
In Section \ref{sec:proofelts} we adapt to open automata several standard properties from refinement relations on LTS.
After that Section \ref{sec:prelref} is dedicated to building refinement relations in the easy cases.
Section \ref{sec:refinement} introduces and analyses the refinement relation on open automata.
Finally Section \ref{sec:ccl} recapitulates the contribution and presents the work left.

\section{State of the art}\label{sec:sota}
Previous work % TODO

@@

An instanciated transition of the automaton is transition produced by any well-formed substitution of the unbound variables of the transition minus the automaton variables.
\end{defi}
The intuition of an open automaton is a partially defined LTS with variables, guards on transitions, and parametrised actions.
The LTS is partial in the sense that one can plug an LTS in the hole and the surrounding LTS will react to the actions emitted by the sub-LTS put in the hole.
Initially the automaton is in the initial state with an extension on all the variables of the initial valuation.
In any configuration, it can perform instanciated transitions which source state is the current state and which guard is satisfiable in the current valuation if the holes emit the indicated actions.
An instanciated transition is a transition where every variable that is not an automaton variable is instantiated with a value.
By performing the instanciated transition, the automaton emits the indicated action and updates its configuration according to the target state and variable assignments.
\begin{noti}[FH-bisimulation]
The FH-bisimulation \cite{henrio:01055091} is noted \(\cong\).

@@

A run of this open automaton can be: The automaton in the hole \(l\) emits many actions, this synchronises with the only transition possible at that moment because \(v = L\).
Then the automaton in hole \(l\) emits a \(\delta\mpar{t}\), this is synchonised with the action that sets \(v \gets R\) and the automata in hole \(r\) must emit a \(\act{acc}\mpar{t}\).
Finally the automaton in the hole \(r\) emits any sequence of actions.

@@

\section{Composition of Open Automata}\label{sec:comp}
Open automata are partially specified automata, that partiality comes mostly from the holes.
The interpretation of a hole is an interface with another open automaton, in which we can plug an open automaton.
The plugging operation is called composition.
In this report, composition will only refer to filling holes, other compositions like ``parallel composition" (pure interleaving) are not tackled (but one can define an open automaton that performs the parallel composition of two others).
The composition of open automata was already implicitly defined by the means of composition on pNets in previous work \cite{henrio:01299562} but never completely formalised on open automata.
The definition of composition below is a direct translation of what happens with pNets composition without the need of introducing pNets.
\begin{defi}[Composition of open automata]
The composition of \(A_c \defobject \OAg[c]\) in the hole \(k \in J_p\) of \(A_p \defobject \OAg[p]\) gives the open automaton defined as follows.
\begin{align*}
	A_p\subst{A_c}{k} \defnotation & \OA{S_p \times S_c}{\mpar{s_{0p}, s_{0c}}}{V_p \uplus V_c}{\sigma_{0p} \uplus \sigma_{0c}}{J_c \uplus J_p \setminus \mbrc{k}}{T} \\
	\text{with } T \defobject & \mset{\OT{\mpar{s_p, s_c}}{\mpar{s'_p, s'_c}}{\alpha_p}{\beta_j^{j \in J'_c \uplus J'_p \setminus \mbrc{k}}}{g_p \wedge g_c \wedge \alpha_c = \beta_k}{\psi_p \uplus \psi_c}}{\OTx{p}{}{}{p} \in T_p, \OTx{c}{}{}{c} \in T_c} \\
	& \cup \mset{\OT{\mpar{s_p, s_c}}{\mpar{s'_p, s_c}}{\alpha_p}{\beta_j^{j \in J'_p}}{g_p}{\psi_p}}{\OTx{p}{}{}{p} \in T_p, k \notin J'_p, s_c \in S_c}
\end{align*}
\end{defi}
The action emitted when \(A_c\) makes a transition is sychronised with the action of the hole \(k\) in transitions of \(A_p\) which have it as a hole action (first transition set, \(\alpha_c = \beta_k\)).
The composition may look like a handshake, with both automata running in parallel (product of states and joint variables) however it is asymmetric because it is the first automaton that decides when the second can evolve, while the first automaton can evolve independently as soon as the hole is not involved in the open transition considered.

@@

\subsection{Properties of the relation on states}
The important properties we consider for the relation on configurations are being a pre-simulation and preventing the introduction of deadlocks.
They are defined in the following.
\begin{defi}[Pre-simulation]
Let \(A_1 \defobject \OAg[1]\) and \(A_2 \defobject \OAg[2]\).
A relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a pre-simulation if it satisfies both
\defitem Initial configurations are related: \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\);
\defitem From related states, all out-transitions from \(A_1\) can be simulated in \(A_2\) and their target states are related:
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}, \forall \sigma: V_1 \uplus V_2 \uplus \fvars{t_1} \to \values, \\
	\mpar{\sigma \vdash R\mpar{s_1, s_2} \wedge g_1} \implies \bigsymb{\exists} t_2 \defobject \OTx{2}{}{2}{2} \in \fOT{s_2}, \exists \nu: \fvars{t_2} \to \values, \\
	\sigma \uplus \nu \vdash \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \wedge g_2 \wedge R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}
\end{multline*}
\end{defi}
This definition translated to open automata the classical definition of simulation \cite{10.1007/3-540-54430-5_78}.
The last formula means that for every pair of related configurations of the automaton and every possible transitions (\(\vdash g_1\)) from the first automaton, there is a possible transition (\(\vdash g_2\)) such that the produced action matches (\(\alpha_1 = \alpha_2\)), the same holes' action (\(J'_1 \cap J'_2\)) matches (\(\beta_{1j} = \beta_{2j}\)) and the target states are related after variable update. % OR variableS
Apart from holes, this is a natural extension of the notion of simulation on LTS, which is the same definition without variables nor holes nor guards.
In FH-bisimulation, holes with the same name will receive the same open automaton when composed, and the resulting open automaton will still be equivalent.
Here we will assume that one of the related open automata can be obtained by composing an open automaton in a hole, open automaton which can have holes itself.
This implies that there is no relation between the related automata's holes, for instance composing \(A_i\) in \(j_i\) (\(1 \leq i \leq n\)) of \(A_0\) gives the set of holes \(\mpar{\biguplus_{0 \leq i \leq n} J_i} \setminus \mset{j_i}{1 \leq i \leq n}\) that is neither subset nor superset of \(J_0\).
Yet we still assume that in holes with the same name, the same open automaton will be composed.
Hence only the holes in common have to match their actions.
However it is not sufficient to imply properties on the relation on automatons like being a preorder.
It is only a necessary condition that we put on any relation that we call a ``simulation".

In addition to the above property, we expect from a refinement relation to prevent the introduction of deadlocks.
Deadlocks are reachable configurations from which no further transition is possible.
They can arise unpredictably when parallelism is involved (which is the case in our composition) and they generally break the intended behaviour of a system.
Preventing them is also a required prerequisite for the refinement to preserve any liveness property.
\begin{defi}[Deadlock prevention, intuitive definition]\label{def:dpid}
A relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is deadlock reducing if:
\defitem It is a pre-simulation;
\defitem If from related states, there is a possible transition in the specification, then there is a pair of possible matching transitions from both automata:
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \forall \sigma: V_1 \uplus V_2 \to \values, \mpar{\sigma \vdash R\mpar{s_1, s_2}} \implies \\
	\mpar{\exists t_2 \in \fOT{s_2}, \exists \nu: \fvars{t_2} \to \values, \sigma \uplus \nu \vdash \fguard{t_2}} \implies \\
	\bigsymb{\exists} \mpar{t_1, t_2} \defobject \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2}, \exists \nu: \fvars{t_1} \uplus \fvars{t_2} \to \values, \\
	\sigma \uplus \nu \vdash g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \wedge R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}
\end{multline*}
\end{defi}
Actually, this requirement prevents the presence of deadlocks in the first automaton if there is a related non-deadlock configuration in the second.
The exact formulation is the contraposition of the one in the previous sentence.
Two reasons motivated the formulation given in Definition \ref{def:dpid}:
\begin{itemize}
\item We require a pair of matching transitions from both automata because we do not want a transition in the implementation making a configuration not a deadlock but this transition cannot be taken from the related configuration of the specification.
\item The alternative formulation of the property would have required the unintuitive quantification on a possibly empty set of transition to state that they are not satisfied.
\end{itemize}
Actually the first point is not necessary, the definition can be simplified into a less intuitive version:
\begin{defi}[Deadlock prevention, working definition]\label{def:dpwd}
A relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is deadlock reducing if it is a pre-simulation and it satisfies the following:
\[ \forall \mpar{s_1, s_2} \in S_1 \times S_2, V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \]
\end{defi}
The requirement here is that from any related configurations, if there is a possible transition in the second automaton (whatever the value of the free variables) then there is a possible transition in the first automaton (the existential quantifier is implicit in the notation \(\vdash\)).
The transition are not matched because the relation on configurations is a pre-simulation and it is sufficient to ensure it as will be shown in the proof.
Apart from the matching transitions this is exactly the same explaination as the intuitive definition.
The expected advantage of this formulation is that a SMT should behave better without the quantifiers.
Indeed the big disjunctions are not quantifiers because in practice the out-transitions are finitary, and so is the expansion of the disjunction.
\begin{lem}\label{lem:dred}
The intuitive and working definitions of deadlock prevention are equivalent.
\end{lem}
Proof of the lemma is given in Appendix \ref{apx:lemdred}.

@@

For transitivity we need to introduce thee open automata.
Let \(A_1 \defobject \OAg[1]\), \(A_2 \defobject \OAg[2]\) and \(A_3 \defobject \OAg[3]\) be open automata.
Let \(R_{12}\) be a hole-\(\triangle\) simulation of \(A_1\) by \(A_2\) and \(R_{23}\) of \(A_2\) by \(A_3\).
The idea for \(R\) is that there exists a state \(s_2 \in S_2\) and values for \(V_2\) such that \(R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3}\).
However we cannot talk about states in formulas.
The solution is to expand the existential quantifier on states as we did in deadlock prevention, which is possible because states are finite.
\[ R \defobject \mpar{s_1, s_3} \mapsto \exists V_2, \bigvee_\subbox{s_2 \in S_2} R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3} \]
\(R\) is a witness of the transitivity, proof is given in \Ludo{Appendix \ref{apx:trans}}{are we supposed to read appendix D before C.2?}.
\end{proofsketch}
\begin{prop}[Limited composition correctness]\label{prop:cc'} Hole-subset simulation is correct w.r.t.\@ composition with automata without holes. \end{prop}
\begin{proofsketch}
For two open automata \(A_1 \defobject \OAg[1]\) and \(A_2 \defobject \OAg[2]\) with \(J_2 = \emptyset\), let \(k \in J_1\) and \(A \defobject A_1\subst{A_2}{k}\).
Let \(\reach{A}\) be a witness that the composition producing \(A\) is non-locking.
The idea for the relation on configurations is that the copy of \(A_1\) in \(A\) do the same things as \(A_1\), as if we were trying to prove reflexivity.
\[ R \defobject \mpar{\mpar{s_{1'}, s_2}, s_1} \mapsto \choice{\reach{A}\mpar{s_{1'}, s_2} \wedge \bigwedge_\subbox{v_1 \in V_1} v_{1'} = v_1 & \text{if } s_{1'} = s_1 \\ \bot & \text{otherwise}} \]
Where \(y_1 \mapsto y_{1'}\) is the renaming for object of \(A_1\) in \(A\).
\(R\) is a witness of the limited composition correctness, proof is given in Appendix \ref{apx:lchs}.
\end{proofsketch}
\begin{prop}[Special composition correctness]
If there is a hole-superset simulation of \(A_1\) by \(A_2\) and \(A_1\subst{A_3}{k}\) is non-locking, then there is a hole-superset simulation of \(A_1\subst{A_3}{k}\) by \(A_2\).
Where \(A_1 \defobject \OAg[1]\), \(A_2 \defobject \OAg[2]\), \(A_3\) are open automata and \(k \in J_1 \setminus J_2\).
\end{prop}
\begin{proofsketch}
Let \(R_{12}\) be a hole-superset simulation of \(A_1\) by \(A_2\) and \(\reach{A_1\subst{A_3}{k}}\) be a witness that \(A_1\subst{A_3}{k}\) is non-locking.
The idea for the relation on configurations is to combine the non-locking composition and the already existing realtion on configurations.
\[ R \defobject \mpar{\mpar{s_1, s_3}, s_2} \mapsto R_{12}\mpar{s_1, s_2} \wedge \reach{A_1\subst{A_3}{k}}\mpar{s_1, s_3} \]
\(R\) is a witness of the special composition correctness, proof is given in Appendix \ref{apx:schs}.
\end{proofsketch}
The hole-subset simulation can only tell if there is a refinement when the composed automata have no holes.
To lift this limitation we need to allow adding holes which doesn't have a direct interpretation.

At first a solution using a function mapping the action performed by the new holes in the composition to the ones performed by removed hole in the base automaton was explored.
It was abandonned because the base automaton has nothing intelligent to tell about the action required by the composed automata.
That is the reason the hole-superset simulation doesn't constrain the action performed by the added holes.
The solution adopted is to give a meaning to adding hole and only adding in a simulation.
With the hole-superset simulation we have an interpretation to added holes:
They are holes imposing a restriction on the transition variables.
However with that solution we cannot remove holes so we cannot express composition corectness, hence the need of the special composition correctness for hole-superset simulation.

For composition correctness, using only one of the hole-\(\triangle\) simulations cannot tell that a non-locking composition with any open automaton is a refinement of the base automaton because we need at the same time to add and remove holes.
If we build the refinement relation as the transitive closure of all the hole-\(\triangle\) simulations then an issue arise.
The transitivity propriety causes that when a hole present in both extrema of the chain but absent in an intermediate automaton, there is a refinement between the extrema of the chain.
In that case all the constraint on holes actions are lifted even if we would like to require an equality of these actions on the transitive closure.

\section{Refinement relation for open automata}\label{sec:refinement}
One of the \Ludo{solutions}{il manque les paragraphes qui presentent le probleme} to keep transitivity and constraints on holes actions from same holes is to keep track of which holes are in common at each step of refinement.
The transitive refinement will have to either rename the holes without constraints or explicitly say which holes have constraints.
\begin{defi}[Open automata refinement]
For two open automata \(A_1 \defobject \OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\) and \(A_2 \defobject \OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\), \(A_1\) is a refinement of \(A_2\) looking at \(H\), noted \(\wrel{A_1}{A_2}{H}\), with \(H \subseteq J_1 \cap J_2\), is defined as:
\begin{multline*}
	\exists R: \mpar{S_1 \times S_2} \to \rformulas[V_1 \uplus V_2], \mpar{\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}} \wedge \forall \mpar{s_1, s_2} \in S_1 \times S_2, \\
	\mpar{\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{\OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X}, \\[12pt]
		\quad \mpar{\forall x \in X, J'_{2x} \cap H = J'_1 \cap H} \\[-10pt]
		\nwedge V_1 \uplus V_2 \uplus \fvars{t_1} \vdash R\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_\subbox{j \in J'_{2x} \cap H} \beta_{1j} = \beta_{2xj} \\
			\nwedge g_{2x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}} \\
	\end{array}} \\
	\wedge V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}
\end{multline*}
\Ludo{major questions}{il me semble qu il y a plein de trucs qu on comprend pas ici: que se passe t il si H est empty? on peut creer un rafinement qui fait n importe quoi avec les trous? on ne perd aucune propriete? peut etrer qu il manque juste un paragraphe qui explique l approche ses points forts et ses limitations mais l j ai pas l intuition}
Giving \(R\) and \(H\) is sufficient to characterise the refinement, so we call \(\mpar{R, H}\) a hole-tracking simulation of \(A_1\) by \(A_2\).
\end{defi}
It is basically the same definition as the refinement relation deduced from hole-\(\triangle\) simulations.
\(R\) is a kind of hole-\(\triangle\) simulation but without the \(\triangle\) constraint on hole sets.
On top of that every actions of the holes outside \(H\) is unconstrained in the related automata.
